---
title: "244 lab in class"
author: "Danny Ettelson"
date: "1/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





```{r libload}


library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car)



```

1. Nonlinear least squares for logistic growth of bacteria

Equation:

Nt = A/(1+B*e^-rt)
B = (K-N0)/N0



```{r}

CellGrowth <- read_csv("CellGrowth.csv")

#Look at it

ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point()+
  theme_classic()

# Find Estimates for logistic growth:

# N = A/(1 + Be^(-rt))

#R will help find a b and r
#Estimate for carrying capacity: 3700
#now if N0 is 2000, what is b? ~0.85
# r ~ 0.1121


B_estimate <- (3700-2000)/2000

#estimate growth rate which is a bit harder b/c exponential

#looks exponential from about t = 0:3 

# Now we're going to estimate r (growth rate) using ggplot

lag_phase <- CellGrowth %>% 
  filter(Time<4)

ggplot(lag_phase, aes(x = Time, y = log(CellCount))) +
  geom_point()

#looks linear! slope of line is estimate of r

lm(log(CellCount) ~ Time , data = lag_phase) # r ~ 0.1121



```

Find model A,B, and r using Gauss-Newton method for iterative convergence (nonlinear least squares):

Use the nls() function:

```{r}

cell_fit <-  nls(CellCount ~ A/(1+ B*exp(-r*Time)), 
                 start = list(A = 3700, B = 0.85, r = 0.1121),
                 data = CellGrowth,
                 trace = TRUE)

#trace shows each iteration's variable estimates and their squared residuals until convergence

cell_fit

#show how model fits data

A <-  coef(cell_fit)[1]
B <-  coef(cell_fit)[2]
r <-  coef(cell_fit)[3]

# Create a sequence of values (times) time_seq to predict with using model

time_seq <- seq(0,20, length = 100)

#Plug that new sequence into my model with the parameters A, B, and r that we found:

cell_pred <- A/(1+ B*exp(-r*time_seq))

#Bind that together with the time sequence data:

pred_df <-  data.frame(time_seq,cell_pred)

#Create a graph with the original data and our model predictions:

ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point( color = "blue", size = 3) +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), color = "orange", size = 1)
  theme_classic()

```

Panel Regression - First Shot

looking at cigarette pack sales data per person since 1985 

State: our entity variable

Year: our time variable

packpc: our dependent variable DV

avgprs: our predictor variable of interest


```{r}

cigs_panel <- Cigarette %>% 
  select(state,year,packpc,avgprs)

ggplot(cigs_panel, aes(x = avgprs, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)+
  theme_classic()

ggplot(cigs_panel, aes(x = year, y = packpc, group = state)) +
  geom_point(aes(color = state)) +
  geom_smooth(method = "lm", aes(color = state), se = FALSE, size = 0.2)+
  theme_classic()

```

Entity fixed effects model:

```{r}

cigs_model <-  plm(packpc ~ avgprs, 
                   data = cigs_panel, 
                   index = c("state","year"),
                   model = "within")

cigs_model

coeftest(cigs_model, vcov. = vcovHC(cigs_model), type = "HC1")



```

Add time fixed effects:

```{r}

et_cigs_model <- plm(packpc ~ avgprs, 
                   data = cigs_panel, 
                   index = c("state","year"),
                   model = "within",
                   effect = "twoways")

et_cigs_model

#test if model has been improved by time effects with pFtest

pFtest(et_cigs_model, cigs_model)

#reject the null! looks as though the model has been improved by adding time effects

```

